---
title: "Coding Assignment on dplyr and ggplot"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
```

#### Load the bodyfat dataset
```{r}
# Load the bodyfat dataset
file = 'Codes/Data/bodyfat.csv'
df = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
str(df)
head(df, n = 2)
```

#### Create lists of categorical and continuous features
```{r}
# Create lists of categorical and continuous features
features = colnames(df)
categorical_features = c('Gender')
continuous_features = features[!(colnames(df) %in% categorical_features)]
```

#### Convert categorical columns to correct type
```{r}
# Convert categorical columns to correct type
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```

#### Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
```{r}
# Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles

df$Weight
mean(df$Weight) # avg
df$Weight-mean(df$Weight) #mean centering rate # deviation vector
(df$Weight-mean(df$Weight))^2 # squared- deviations vector
 # mean of sqaure deviation 
mean((df$Weight-mean(df$Weight))^2) # sample variance
var(df$Weight) # sample variance using in-built function 
sqrt(mean((df$Weight-mean(df$Weight))^2)) # sample standard deviations
sd(df$Weight) # sample standard deviations using in-built function
(df$Weight -mean(df$Weight))/sd(df$Weight) # standardizes weight values with no units (bcz kg kg cancels)  ## Also called Z-scores of weights

scale(df$Weight) # standardizes weight values USING WEIGHT VALUES

## weights in pounds, df['Weightlbs']=
## weights in grams , df['Weightgms']=
#df$WeightgmsS[1] op-70 42. de$Weightlbs[1]  op=157.74, gma[1]   op=70420
```

#### Plot the distribution of males and females in the dataset
```{r}
# Plot the distribution of males and females in the dataset
p = ggplot(data = df) +
  geom_bar(aes(x=Gender),color= 'black',fill='steel blue') +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Distribution of Genders')
p+coord_flip()
#p

```

# continuous variable 
#### Plot a component plot of weight values ##component plot-- plot with respect to values of single continuous variable here is weight
```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Weight)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot ofWeights')
p
```

#get red line
```{r}
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Weight)) +
  geom_hline(yintercept = mean(df$Weight), color='red')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```

# get dotted/dashed red line
```{r}
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Weight)) +
  geom_hline(yintercept = mean(df$Weight), color='red', linetype='dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```


# better plot # bcz -ve and +ve points 
```{r}
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Weight-mean(Weight))) +
  ##geom_hline(yintercept = mean(df$Weight), color='red', linetype='dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Mean-centered Weights')
p
```

#### Plot a component plot of Bodyfat values
```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Bodyfat (g)',
       title = 'Component Plot of Bodyfat')
p
```


```{r}
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Bodyfat)) +
  geom_hline(yintercept = mean(df$Bodyfat), color='red', linetype='dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Bodyfat (g)',
       title = 'Component Plot of Bodyfat')
p
```


```{r}
p = ggplot(data = df) +
  geom_point(aes(x=1:nrow(df),y=Bodyfat-mean(Bodyfat))) +
  ##geom_hline(yintercept = mean(df$Weight), color='red', linetype='dashed')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Bodyfat (g)',
       title = 'Component Plot of Mean-centered Bodyfat')
p
```

# histogram
#Plot a histogram of weight values
```{r}
# Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x=Weight, y=after_stat(count)), breaks=seq(mean_weight-3*sd_weight, mean_weight+3*sd_weight, by=2.0),color='black', fill='steel blue',alpha = 0.4)+ 
  geom_vline(aes(xintercept = mean(Weight)), color='red',linetype="dashed", size=1)+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a histogram of weight values
```{r}

## don't refer
# Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x=Weight, y=after_stat(count()))) +
  #geom_vline(?)+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a histogram of height values
```{r}
# Plot a histogram of height values
mean_height = mean(df$Height)
sd_height = sd(df$Height)
p = ggplot(data = df) +
  geom_histogram(aes(x=Height, y=after_stat(count)), breaks=seq(mean_height-3*sd_height, mean_height+3*sd_height, by=0.5),color='black', fill='steel blue',alpha = 0.4)+ 
  geom_vline(aes(xintercept = mean(Height)), color='red',linetype="dashed", size=1)+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Height (inch)',
       y = 'Frequency',
       title = 'Frequency Histogram of Heights')
p

```

#### Plot a histogram of bodyfat values
```{r}
# Plot a histogram of bodyfat values
mean_bodyfat = mean(df$Bodyfat)
sd_bodyfat = sd(df$Bodyfat)
p = ggplot(data = df) +
  geom_histogram(aes(x=Bodyfat, y=after_stat(count)), breaks=seq(mean_bodyfat-3*sd_bodyfat, mean_bodyfat+3*sd_bodyfat, by=1.0),color='black', fill='steel blue',alpha = 0.4)+ 
  geom_vline(aes(xintercept = mean(Bodyfat)), color='red',linetype="dashed", size=1)+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Bodyfat (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Bodyfats')
p
```

#### Plot a histogram of weights for both genders
```{r}
# Plot a histogram of weights for both genders
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x=Weight, y=after_stat(count),color=Gender))+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights for male and female')
p

```

#### Plot a boxplot of bodyfat values
```{r}
# Plot a boxplot of bodyfat values

p = ggplot(data = df) +
  geom_boxplot(aes(y=Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Frequency Boxplot of Weights for male and female')
p
```

#### Plot a boxplot of bodyfat values for both genders
```{r}
# Plot a boxplot of bodyfat values for both genders


p = ggplot(data = df) +
  geom_boxplot(aes(x=Gender, y=Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Boxplot of Frequency bodyfat of Weights for male and female')
p

```

#### Introduce a new categorical column in the dataframe called 'Weightlevel' with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
```{r}
# Introduce a new categorical column in the dataframe called weight level with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)


#df['Wlevel']=df['Weight'] %>% mutate(Wlevel = ifelse(Weight < quantile(Weight, 0.25), "low",
                       #ifelse(Weight > quantile(Weight, 0.75), "high", "medium")))
#xgender y=weight levels


#df['Wlevel']=df['Weight'] %>% mutate(Weightlevel = case_when(
    #Weight < quantile(Weight, 0.25) ~ "low",
   # Weight <= (quantile(Weight, 0.25) + IQR(Weight)) ~ "medium"))




q25=quantile(df$Weight,0.25)

q75=quantile(df$Weight,0.75)

df=df %>% mutate(Weightlevel = case_when(Weight<=q25~'low', Weight>q25 & Weight<q75~'Medium',Weight>q75~'High' ))
```


```{r}
categorical_cols=c('Weightlevel')
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```
#### Convert categorical columns to correct type
# Convert categorical columns to correct type
```{r}
p = ggplot(data = df) +

  geom_boxplot(aes(x=Gender,y=Bodyfat, fill=Weightlevel))+

  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),

  axis.text = element_text(size = 12),

  axis.text.x = element_text(size = 14),

  axis.text.y = element_text(size = 14),

  axis.title = element_text(size = 14, face = "bold"))+

  labs(x = 'Weightlevel',

       y = 'Bodyfat',

       title = 'boxplot of bodyfat values for both weightlevel')

p
 



```


```{r}
#df <- df %>% select(-Wlevel)

```


#### Plot a boxplot of bodyfat values for both genders across the three Weightlevel groups
```{r}
# Plot a boxplot of bodyfat values for both genders both genders across the three Weightlevel groups
````

#### The 68-95-99 rule
```{r}
# The 68-95-99 rule

```


