---
title: "Intro to R"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Load the food texture dataset
```{r}
file = 'Codes/Data/food-texture.csv'
#file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(foodData, 2)
str(foodData)
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispylevel = ifelse(Crispy > 11, 'High', 'Low'))
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy (8 levels 8 through 15) and Crispylevel (2 levels 0 and 1)
# Crispy is a categorical feature with an order
str(foodData)
categorical_cols = c('Crispy', 'Crispylevel')
foodData[categorical_cols] = lapply(foodData[categorical_cols], as.factor)
str(foodData)
```

#### Visualize the Oil Percentage feature (continuous) using a histogram
```{r}
p = ggplot(data = foodData) +
  geom_histogram(aes(x = Oil), binwidth = 1, color = 'steelblue')
p
```

```{r}
p = ggplot(data = foodData[foodData$Crispy == 13, ]) +
  geom_histogram(aes(x = Density), binwidth = 100, color = 'blue')
p
```

```{r}
p = ggplot(data = foodData) +
  geom_histogram(aes(x = Hardness), binwidth = 10, color = 'steelblue')
p
```



#### Scatter plot between Oil Percentage and Density (relating two continuous variables)
```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil), color = 'red') +
  labs(x = 'Density', y = 'Oil percentage', title = 'Pastry Analysis')
p
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil)) +
  labs(x = 'Density', y = 'Oil percentage', title = 'Pastry Analysis')
p
```

```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil, color = Crispy, shape= Crispylevel)) +
  labs(x = 'Density', y = 'Oil percentage', title = 'Pastry Analysis')
p

```


```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil, color = Crispylevel)) +
  labs(x = 'Density', y = 'Oil percentage', title = 'Pastry Analysis')
p

```


#### Scatter plot between Density and Hardness
```{r}
```

#### Scatter plot between Density and Fracture
```{r}
```

#### Smooth line plot using ggplot
```{r}
p = ggplot(data = foodData) +
  geom_smooth(aes(x = Oil, y = Density))
p
```

#### Barplot for Crispy (categorical variable)
```{r}
p = ggplot(data = foodData) +
  geom_bar(aes(x = Crispy))
p
```

#### In-built functions for dataframes
```{r}
# Mean oil percentage across all samples

# Mean-centering of OilPercentage

# Sum of the squared deviation from the mean

# Average of the squared deviation from the mean

# Variance of OilPercentage

# Standard deviation of OilPercentage
```


#### Box plot using ggplot with color coding
```{r}
p = ggplot(data = foodData) +
  geom_boxplot(aes(y = Oil))
p
```

```{r}
p = ggplot(data = foodData) +
  geom_boxplot(aes(x = Crispylevel, y = Oil))
p
```

```{r}
p = ggplot(data = foodData[foodData$Crispylevel == 'Low', ]) +
  geom_histogram(aes(x = Density), binwidth = 100, color = 'blue')
p
```

#####percentile and quantiles

# what is a percentile? sometimes also reffered to 
# as the quantile
```{r}
foodData %>% select(Oil)
median(foodData$Oil) # 50 th percentile or 0.5th quantile
mean(foodData$Oil)
```

```{r}
foodData %>% select(Oil)
Oil_median=median(foodData$Oil) # 50 th percentile or 0.5th quantile
#mean(foodData$Oil)
print(Oil_median)
```

# find the fraction of samples whose oil percentage is less than or equal to median oil percentage (50th percentile)=0.5

#P(Random sample's oil percentage <=median oil percentage)= 0.5
```{r}
foodData %>% select(Oil)
Oil_median=median(foodData$Oil) # 50 th percentile or 0.5th quantile
#mean(foodData$Oil)
print(Oil_median)

#foodData %>% filter(Oil <= Oil_median) %>% select(Oil)
nrow(foodData %>% filter(Oil <= Oil_median) %>% select(Oil))/nrow(foodData)
#foodData$Oil --> vector output
#foodData$Oil <=Oil_median --> true / false o/p
mean(foodData$Oil <=Oil_median)
```


#Diabetes

#### Load the food Diabetes dataset
```{r}
file = 'Codes/Data/diabetes.csv'
#file = 'Data/food-texture.csv'
diabetData = read.csv(file, header = TRUE, stringsAsFactors = FALSE)
head(diabetData, 2)
str(diabetData)
```

```{r}
categorical_cols=c('Outcome')
foodData[categorical_cols] = lapply(diabetData[categorical_cols], as.factor)
str(foodData)
```


```{r}
#diabetData = diabetData %>% mutate(Plevel = ifelse(Pregnancies > 10, 'High', 'Low'))
```

```{r}
str(diabetData)
categorical_cols = c('Pregnancies', 'Outcome')
diabetData[categorical_cols] = lapply(diabetData[categorical_cols], as.factor)
str(diabetData)
```


#### Visualize the Glucose feature (continuous) using a histogram
```{r}
p = ggplot(data = diabetData) +
  geom_histogram(aes(x = Glucose), binwidth = 10, color = 'steelblue')
p
```

```{r}
mean(diabetData$Glucose)
median(diabetData$Glucose)
mean(diabetData$Glucose <= median(diabetData$Glucose))
sum(diabetData$Glucose <= median(diabetData$Glucose))
nrow(diabetData)

quantile(diabetData$Glucose,0.5) # P(Glucose <+0.5th quantile)=0.5
quantile(diabetData$Glucose,0.25)
sum(diabetData$Glucose<=99)/nrow((diabetData))
```

#### Scatter plot between Glucose and Insulin (relating two continuous variables)
```{r}
p = ggplot(data = diabetData) +
  geom_point(aes(y = Glucose, x = Insulin), color = 'red') +
  labs(y = 'Glucose', x = 'Insulin', title = 'Diabetes Analysis')
p
```


```{r}
p = ggplot(data = diabetData) +
  geom_point(aes(y = Glucose, x = Insulin, color = Outcome)) +
  labs(y = 'Glucose', x = 'Insulin', title = 'Diabetes Analysis')
p

```


```{r}
data = diabetData %>% filter(Insulin!=0)
data
```

```{r}
p = ggplot(data = diabetData %>% filter(Insulin!=0)) +
  geom_point(aes(y = Glucose, x = Insulin, color = Outcome)) +
  labs(y = 'Glucose', x = 'Insulin', title = 'Diabetes Analysis')
p

```


```{r}
diabetData %>% select(Glucose)
Glucose_median=median(diabetData$Glucose) # 50 th percentile or 0.5th quantile
print(Glucose_median)
nrow(diabetData %>% filter(Glucose <= Glucose_median) %>% select(Glucose))/nrow(diabetData)
mean(diabetData$Glucose <=Glucose_median)
```



```{r}
p = ggplot(data = diabetData %>% filter(Insulin!=0)) +
  geom_point(aes(x = Age, y = Glucose, color = Outcome)) +
  labs(y = 'Glucose', x = 'Age', title = 'Diabetes Analysis')
p

```


```{r}
p = ggplot(data = diabetData) +
  geom_boxplot(aes(x = Outcome, y = Glucose))
p
```

```{r}
p = ggplot(data = diabetData) +
  geom_boxplot(aes(y = Glucose))
p
```


```{r}
p = ggplot(data = diabetData) +
  geom_boxplot(aes(y = Age))
p
```

```{r}
p = ggplot(data = diabetData) +
  geom_histogram(aes(x = Age), binwidth = 10, color = 'steelblue')
p
```

```{r}
mean(diabetData$Age)
median(diabetData$Age)
```


```{r}
p = ggplot(data = diabetData) +
  geom_boxplot(aes(y = Age))
p
```

```{r}
p = ggplot(data = diabetData) +
  geom_boxplot(aes(y = Age , x= Outcome))
p
```

```{r}
#p = ggplot(data = diabetData) +
  #geom_boxplot(aes(y = Outcome))
#p
```

```{r}
p = ggplot(data = diabetData  %>% filter(Insulin!=0)) +
  geom_boxplot(aes(y = Age , x= Outcome))
p
```

```{r}

```

```{r}
```
